### 2020-09-27 priority-cp.tex
Updated text.
Added text to PGV. Outlined lemmas for PGV closed progress.

In this section we present an updated version of Priority CP~\cite{dardha2018}, which is Wadler's Classical Processes~\cite{wadler2012} with \emph{priorities}. Some details of the calculus are inspired by Carbone \emph{et al.}~\cite{carbone2016} and Caires and P\'{e}rez~\cite{caires2017}.

\paragraph*{Priorities, types, and environments}
Types are based on classical linear logic formulas, and are annotated with priorities $\cs{o}\in\mathbb{N}$.

The types $\ty{\tytens[\cs{o}]{A}{B}}$ and $\ty{\typarr[\cs{o}]{A}{B}}$ type the endpoints of a channel over which we send or receive a channel of type $\ty{A}$, and then proceed as $\ty{B}$.
The types $\ty{\tyone[\cs{o}]}$ and $\ty{\tybot[\cs{o}]}$ type the endpoints of a channel over which we send or receive a ping, and then close the channel. These act as the units for $\ty{\tytens[\cs{o}]{A}{B}}$ and $\ty{\typarr[\cs{o}]{A}{B}}$.

>>>>

We present an updated version of Priority CP~\cite[PCP]{dardha2018}, which is Wadler's Classical Processes~\cite[CP]{wadler2012} with \emph{priorities}.
The main change is the move away from a reduction strategy based on cut elimination using commuting conversions, and towards a reduction strategy using the structural congruence.

\paragraph*{Priorities, types, and environments}
Types are based on classical linear logic connectives, and are annotated with priorities $\cs{o}\in\mathbb{N}$.

The types $\ty{\tytens[\cs{o}]{A}{B}}$ and $\ty{\typarr[\cs{o}]{A}{B}}$ type the endpoints of a channel over which we send or receive a channel of type $\ty{A}$, and then proceed as $\ty{B}$.
The types $\ty{\tyone[\cs{o}]}$ and $\ty{\tybot[\cs{o}]}$ type the endpoints of a channel whose session has finished, and over which we send or receive a ping before closing the channel. These act as the units for $\ty{\tytens[\cs{o}]{A}{B}}$ and $\ty{\typarr[\cs{o}]{A}{B}}$.

### 2020-09-27 priority-gv.tex
Added text to PGV. Outlined lemmas for PGV closed progress.

\paragraph*{Priorities, types, and environments}
\[
  \cs{o}, \cs{p}, \cs{q}, \cs{r} \in \mathbb{N}\cup\{\cs{\pbot},\cs{\ptop}\}
\]
\[
\begin{array}{lcl}
  \ty{S}
  & \coloneqq & \ty{\tysend[\cs{o}]{T}{S}}
    \sep        \ty{\tyrecv[\cs{o}]{T}{S}}
    \sep        \ty{\tyends[\cs{o}]}
    \sep        \ty{\tyendr[\cs{o}]}
  \\
  \ty{T}, \ty{U}
  & \coloneqq & \ty{\tyunit}
    \sep        \ty{\tyvoid}
    \sep        \ty{\typrod{T}{U}}
    \sep        \ty{\tysum{T}{U}}
    \sep        \ty{\tylolli[\cs{p},\cs{q}]{T}{U}}
    \sep        \ty{S}
  \\
  \ty{\Gamma}, \ty{\Delta}
  & \coloneqq & \ty{\emptyenv}
    \sep        \ty{\Gamma}, \tmty{x}{T}
\end{array}
\]
\[
\begin{array}{lcl}
  \ty{\co{\tysend[\cs{o}]{T}{S}}} & = & \ty{\tyrecv[\cs{o}]{T}{\co{S}}} \\
  \ty{\co{\tyrecv[\cs{o}]{T}{S}}} & = & \ty{\tysend[\cs{o}]{T}{\co{S}}}
\end{array}
\qquad
\begin{array}{lcl}
  \ty{\co{\tyends[\cs{o}]}} & = & \ty{\co{\tyendr[\cs{o}]}} \\
  \ty{\co{\tyendr[\cs{o}]}} & = & \ty{\co{\tyends[\cs{o}]}}
\end{array}
\]
\[
\begin{array}{lcl}
  \pr(\ty{\tysend[\cs{o}]{T}{S}}) & = & \cs{o} \\
  \pr(\ty{\tyrecv[\cs{o}]{T}{S}}) & = & \cs{o} \\
  \pr(\ty{\typrod{T}{U}}) & = & \pr({\ty{T}})\sqcap\pr({\ty{U}}) \\
  \pr(\ty{\tysum{T}{U}})  & = & \pr({\ty{T}})\sqcap\pr({\ty{U}}) \\
  \pr(\ty{\tylolli[\cs{p},\cs{q}]{T}{U}}) & = & \cs{p} \\
  \pr(\ty{\Gamma}, \tmty{x}{A}) & = & \pr(\ty{\Gamma})\sqcap\pr(\ty{A})
\end{array}
\qquad
\begin{array}{lcl}
  \pr(\ty{\tyends[\cs{o}]}) & = & \cs{o} \\
  \pr(\ty{\tyendr[\cs{o}]}) & = & \cs{o} \\
  \pr(\ty{\tyunit})         & = & \ptop \\
  \pr(\ty{\tyvoid})         & = & \ptop \\
  \pr(\ty{S})               & = & \pr(\ty{S}) \\
  \pr(\ty{\emptyenv})       & = & \ptop
\end{array}
\]

\paragraph*{Syntactic Sugar for Types}
\[
  \ty{\tylolli{T}{U}} \elabarrow \ty{\tylolli[\cs{\ptop},\cs{\pbot}]{T}{U}}
\]

@@

\paragraph*{Runtime Term Syntax}
\[
\begin{array}[t]{lcl}
  \tm{\conf{C}}, \tm{\conf{D}}, \tm{\conf{E}}
  & \coloneqq & \tm{\phi\;M}
    \sep        \tm{\ppar{\conf{C}}{\conf{D}}}
    \sep        \tm{\res{x}{y}{\conf{C}}}
  \\
  \tm{\phi}
  & \coloneqq & \tm{\main}
    \sep        \tm{\child}
  \\
  \tm{V}, \tm{W}
  & \coloneqq & \tm{x}
    \sep        \tm{K}
    \sep        \tm{\lambda x.M}
    \sep        \tm{\unit}
    \sep        \tm{\pair{M}{N}}
    \sep        \tm{\inl{M}}
    \sep        \tm{\inr{M}}
  \\
  \tm{R}
  & \coloneqq & \tm{K\;V}
  \\
  \tm{E}
  & \coloneqq & \tm{\hole} \\
  & \sep      & \tm{E\;M}
    \sep        \tm{V\;E} \\
  & \sep      & \tm{\letunit{E}{N}} \\
  & \sep      & \tm{\pair{E}{M}}
    \sep        \tm{\pair{V}{E}}
    \sep        \tm{\letpair{x}{y}{E}{M}} \\
  & \sep      & \tm{\inl{E}}
    \sep        \tm{\inr{E}}
    \sep        \tm{\casesum{E}{x}{M}{y}{N}} \\
  & \sep      & \tm{\absurd{E}} \\
  \\
  \tm{\conf{F}}
  & \coloneqq & \tm{\phi\;E}
  \\
  \tm{\conf{G}}
  & \coloneqq & \tm{\hole} \\
  & \sep      & \tm{\ppar{\conf{G}}{\conf{C}}}
    \sep        \tm{\res{x}{y}{\conf{G}}}
\end{array}
\]
\[
\begin{array}{ll}
  \tm{\main}  + \tm{\child} = \tm{\main}
  \\
  \tm{\child} + \tm{\main}  = \tm{\main}
\end{array}
\quad
\begin{array}{ll}
  \tm{\child} + \tm{\child} = \tm{\child}
  \\
  \tm{\main}  + \tm{\main} \; \text{undefined}
\end{array}
\]

@@

\begin{restatablelemma}{lempgvvaluedone}
  \label{lem:pgv-value-done}
  If $\tseq[\cs{o}]{\ty{\Gamma}}{V}{T}$, then $\cs{o}=\cs{\pbot}$, and $\pr(\ty{\Gamma})=\pr(\ty{T})$.
\end{restatablelemma}
\begin{proof}
  By induction on the derivation of $\tseq[\cs{o}]{\ty{\Gamma}}{V}{T}$.
  (Details in \cref{prf:lem-pgv-value-done}.)
\end{proof}

@@

\begin{restatablelemma}{lempgvsubjectreductionterms}[Subject Reduction, $\tred$]
  \label{lem:pgv-subject-reduction-terms}
  \hfill\\%newline before theorem statement
  If $\tseq[\cs{o}]{\ty{\Gamma}}{M}{T}$ and $\tm{M}\tred\tm{M'}$,
  then $\tseq[\cs{o}]{\ty{\Gamma}}{M'}{T}$.
\end{restatablelemma}
\begin{proof}
  By induction on the derivation of $\tm{M}\tred\tm{M'}$.
  (Details in \cref{prf:lem-pgv-subject-reduction-terms}.)
\end{proof}

@@

>>>>

\paragraph*{Priorities, types, and environments}
Types are divided into \emph{session types} $\ty{S}$ and \emph{value types} $\ty{T}$, $\ty{U}$.

Session types are annotated with \emph{priorities} $\cs{o}\in\mathbb{N}$.
The types $\ty{\tysend[\cs{o}]{T}{S}}$ and $\ty{\tyrecv[\cs{o}]{T}{S}}$ type the endpoints of a channel over which we send or receive a value of type $\ty{T}$, and then proceed as $\ty{S}$.
The types $\ty{\tyends[\cs{o}]}$ and $\ty{\tyendr[\cs{o}]}$ type endpoints of a channel whose session has finished, and over which we send or receive a ping before closing the channel.
\[
\begin{array}{lcl}
  \ty{S}
  & \coloneqq & \ty{\tysend[\cs{o}]{T}{S}}
    \sep        \ty{\tyrecv[\cs{o}]{T}{S}}
    \sep        \ty{\tyends[\cs{o}]}
    \sep        \ty{\tyendr[\cs{o}]}
\end{array}
\]
Value types are based on the types of the linear $\lambda$-calculus.
The types $\ty{\typrod{T}{U}}$, $\ty{\tyunit}$, $\ty{\tysum{T}{U}}$, and $\ty{\tyvoid}$ are the standard linear $\lambda$-calculus product type, unit type, sum type, and empty type.
The type $\ty{\tylolli[\cs{p},\cs{q}]{T}{U}}$ is the standard linear function type, annotated with \emph{priority bounds} $\cs{p},\cs{q}\in\mathbb{N}\cup\{\cs{\pbot},\cs{\ptop}\}$.
Session types are value types, as endpoints are values.
\[
\begin{array}{lcl}
  \ty{T}, \ty{U}
  & \coloneqq & \ty{\typrod{T}{U}}
    \sep        \ty{\tyunit}
    \sep        \ty{\tysum{T}{U}}
    \sep        \ty{\tyvoid}
    \sep        \ty{\tylolli[\cs{p},\cs{q}]{T}{U}}
    \sep        \ty{S}
\end{array}
\]
Given a function with type $\ty{\tylolli[\cs{p},\cs{q}]{T}{U}}$, the priority bound $\cs{p}$ is a \emph{lower bound} on the priorities of the channel names captured by the body of the function, and the priority bound $\cs{q}$ is an \emph{upper bound} on the priority of the communications that take place as a result of applying the function.
We discuss this in more detail when we give the typing rules.
We define syntactic sugar for the type of \emph{pure functions}.
\[
  \ty{\tylolli{T}{U}} \elabarrow \ty{\tylolli[\cs{\ptop},\cs{\pbot}]{T}{U}}
\]
A typing context $\ty\Gamma$ is a partial function associating names to types.
\[
\begin{array}{lcl}
  \ty{\Gamma}, \ty{\Delta}
  & \coloneqq & \ty{\emptyenv}
    \sep        \ty{\Gamma}, \tmty{x}{T}
\end{array}
\]
\[
\begin{array}{lcl}
  \ty{\co{\tysend[\cs{o}]{T}{S}}} & = & \ty{\tyrecv[\cs{o}]{T}{\co{S}}} \\
  \ty{\co{\tyrecv[\cs{o}]{T}{S}}} & = & \ty{\tysend[\cs{o}]{T}{\co{S}}}
\end{array}
\qquad
\begin{array}{lcl}
  \ty{\co{\tyends[\cs{o}]}} & = & \ty{\co{\tyendr[\cs{o}]}} \\
  \ty{\co{\tyendr[\cs{o}]}} & = & \ty{\co{\tyends[\cs{o}]}}
\end{array}
\]
\[
\begin{array}{lcl}
  \pr(\ty{\tysend[\cs{o}]{T}{S}})            & = & \cs{o} \\
  \pr(\ty{\tyrecv[\cs{o}]{T}{S}})            & = & \cs{o} \\
\end{array}
\qquad
\begin{array}{lcl}
  \pr(\ty{\tyends[\cs{o}]})                  & = & \cs{o} \\
  \pr(\ty{\tyendr[\cs{o}]})                  & = & \cs{o} \\
\end{array}
\]
\[
\begin{array}{lcl}
  \minpr(\ty{\typrod{T}{U}})                 & = & \minpr({\ty{T}})\sqcap\minpr({\ty{U}}) \\
  \minpr(\ty{\tysum{T}{U}})                  & = & \minpr({\ty{T}})\sqcap\minpr({\ty{U}}) \\
  \minpr(\ty{\tylolli[\cs{p},\cs{q}]{T}{U}}) & = & \cs{p} \\
  \\
  \minpr(\ty{\Gamma}, \tmty{x}{A})           & = & \minpr(\ty{\Gamma})\sqcap\pr(\ty{A})
\end{array}
\qquad
\begin{array}{lcl}
  \minpr(\ty{\tyunit})                       & = & \ptop \\
  \minpr(\ty{\tyvoid})                       & = & \ptop \\
  \minpr(\ty{S})                             & = & \pr(\ty{S}) \\
  \\
  \minpr(\ty{\emptyenv})                     & = & \ptop
\end{array}
\]

@@

\paragraph*{Runtime Term Syntax}
\[
\begin{array}[t]{lcl}
  \tm{\conf{C}}, \tm{\conf{D}}, \tm{\conf{E}}
  & \coloneqq & \tm{\phi\;M}
    \sep        \tm{\ppar{\conf{C}}{\conf{D}}}
    \sep        \tm{\res{x}{y}{\conf{C}}}
  \\
  \tm{\phi}
  & \coloneqq & \tm{\main}
    \sep        \tm{\child}
  \\
  \tm{V}, \tm{W}
  & \coloneqq & \tm{x}
    \sep        \tm{K}
    \sep        \tm{\lambda x.M}
    \sep        \tm{\unit}
    \sep        \tm{\pair{M}{N}}
    \sep        \tm{\inl{M}}
    \sep        \tm{\inr{M}}
  \\
  \tm{E}
  & \coloneqq & \tm{\hole} \\
  & \sep      & \tm{E\;M}
    \sep        \tm{V\;E} \\
  & \sep      & \tm{\letunit{E}{N}} \\
  & \sep      & \tm{\pair{E}{M}}
    \sep        \tm{\pair{V}{E}}
    \sep        \tm{\letpair{x}{y}{E}{M}} \\
  & \sep      & \tm{\inl{E}}
    \sep        \tm{\inr{E}}
    \sep        \tm{\casesum{E}{x}{M}{y}{N}} \\
  & \sep      & \tm{\absurd{E}} \\
  \\
  \tm{\conf{F}}
  & \coloneqq & \tm{\phi\;E}
  \\
  \tm{\conf{G}}
  & \coloneqq & \tm{\hole} \\
  & \sep      & \tm{\ppar{\conf{G}}{\conf{C}}}
    \sep        \tm{\res{x}{y}{\conf{G}}}
\end{array}
\]
\[
\begin{array}{ll}
  \tm{\main}  + \tm{\child} = \tm{\main}
  \\
  \tm{\child} + \tm{\main}  = \tm{\main}
\end{array}
\quad
\begin{array}{ll}
  \tm{\child} + \tm{\child} = \tm{\child}
  \\
  \tm{\main}  + \tm{\main} \; \text{undefined}
\end{array}
\]

@@

\begin{restatablelemma}{lempgvvaluedone}
  \label{lem:pgv-value-done}
  If $\tseq[\cs{p}]{\ty{\Gamma}}{V}{T}$, then $\cs{p}=\cs{\pbot}$, and $\minpr(\ty{\Gamma})=\minpr(\ty{T})$.
\end{restatablelemma}
\begin{proof}
  By induction on the derivation of $\tseq[\cs{p}]{\ty{\Gamma}}{V}{T}$.
  (Details in \cref{prf:lem-pgv-value-done}.)
\end{proof}

@@

\begin{restatablelemma}{lempgvsubjectreductionterms}[Subject Reduction, $\tred$]
  \label{lem:pgv-subject-reduction-terms}
  \hfill\\%newline before theorem statement
  If $\tseq[\cs{p}]{\ty{\Gamma}}{M}{T}$ and $\tm{M}\tred\tm{M'}$,
  then $\tseq[\cs{p}]{\ty{\Gamma}}{M'}{T}$.
\end{restatablelemma}
\begin{proof}
  By induction on the derivation of $\tm{M}\tred\tm{M'}$.
  (Details in \cref{prf:lem-pgv-subject-reduction-terms}.)
\end{proof}

@@

\begin{definition}[Actions]
  \label{def:pgv-actions}
  A~term acts on an endpoint $\tm{x}$ if it is in one of the following forms:
  \begin{multicols}{3}
    \begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
    \item $\tm{\link\;\pair{x}{y}}$ 
    \item $\tm{\link\;\pair{y}{x}}$
    \item $\tm{\send\;\pair{x}{y}}$
    \item $\tm{\recv\;{x}}$
    \item $\tm{\close\;{x}}$
    \item $\tm{\wait\;{x}}$
    \end{itemize}
  \end{multicols}
  \noindent
  A~term is an action if it acts on some endpoint $\tm{x}$.
\end{definition}

\begin{definition}[Ready Actions]
  \label{def:pgv-ready-actions}
  A~term is ready to act on an endpoint $\tm{x}$ if it is of the form $\tm{\plug{E}{M}}$, where $\tm{M}$ acts on $\tm{x}$. A~term is ready if it is ready to act on some endpoint $\tm{x}$.
\end{definition}

\begin{restatablelemma}{lempgvprogressterms}[Open Progress, $\tred$]

\end{restatablelemma}

\begin{definition}[Canonical Forms]
  \label{def:pgv-canonical-forms}
  A~configuration $\tm{\conf{C}}$ is in canonical form if it is of the following form:
  \[
    \tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(\child\;M_1\parallel\dots\parallel\child\;M_m\parallel\main\;N)}}}
  \]
\end{definition}

\begin{restatablelemma}{lempgvcanonicalforms}[Canonical Forms]
  \label{lem:pgv-canonical-forms}
  If $\cseq[\main]{\ty{\Gamma}}{\conf{C}}$, there exists some $\tm{\conf{D}}$ such that $\tm{\conf{C}}\equiv\tm{\conf{D}}$ and $\tm{\conf{D}}$ is in canonical form.
\end{restatablelemma}
\begin{proof}
  We move any $\nu$-binders to the top using \LabTirName{SC-ResExt}, discard any superfluous occurrences of $\tm{\child\;\unit}$ using \LabTirName{SC-ParNil}, and move the main thread to the rightmost position using \LabTirName{SC-ParComm} and \LabTirName{SC-ParAssoc}.
\end{proof}

\begin{restatabletheorem}{thmpgvclosedprogressconfs}[Closed Progress, $\cred$]
  \label{lem:pgv-closed-progress-confs}
  If $\cseq[\main]{\emptyenv}{\conf{C}}$ and $\tm{\conf{C}}$ is in canonical form, then either $\tm{\conf{C}}=\tm{\main\;V}$ or there exists a $\tm{\conf{D}}$ such that $\tm{\conf{C}}\cred\tm{\conf{D}}$.
\end{restatabletheorem}
\begin{proof}
  \label{prf:thm-pgv-closed-progress}
  Let $\tm{\conf{C}}=\tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(\child\;M_1\parallel\dots\parallel\child\;M_m\parallel\main\;N)}}}$. If any term $\tm{M_i}$ or $\tm{N}$ is \emph{not} ready, 
\end{proof}

  % Pick the process $\tm{P_i}$ with the smallest priority $\minpr{(\ty{\Gamma_i})}$. $\tm{P_i}$ acts on some endpoint $\tmty{y}{A}\in\ty{\Gamma_i}$. We must have $\minpr{(\ty{\Gamma_i})}=\pr{(\ty{A})}$, since the other actions in $\tm{P_i}$ are guarded by the action on $\tmty{y}{A}$.

  % If $\tm{P_j}$ is a link $\tm{\link{y}{z}}$ or $\tm{\link{z}{y}}$, we apply \LabTirName{E-Link}.

  % Otherwise, there must be a process $\seq{P_j}{\ty{\Gamma_j}}$ where $\tmty{y'}{\co{A}}\in\ty{\Gamma_j}$ and $\tm{y}$ and $\tm{y'}$ are dual endpoints of the same channel, since $\tm{y}$ is bound. By duality, $\pr{(\ty{A})}=\pr{(\ty{\co{A}})}$. We must have $\pr{(\ty{\co{A}})}=\minpr{(\ty{\Gamma_j})}$, since we chose $\tm{P_i}$ to have the smallest priority $\minpr{(\Gamma_i)}$. We cannot have $\tm{P_i}=\tm{P_j}$, otherwise the action on $\tm{y'}$ would be guarded by the action on $\tm{y}$, requiring $\pr{(\ty{A})}<\pr{(\ty{\co{A}})}$. The process $\tm{P_j}$ must act on $\tm{y'}$, otherwise the action on $\tm{y'}$ would be guarded by another action with priority smaller than $\pr{(\ty{A})}$. We have two processes, acting on dual endpoints. We apply the appropriate reduction rule, \ie \LabTirName{E-Send}, \LabTirName{E-Close}, \LabTirName{E-Select-Inl}, or \LabTirName{E-Select-Inr}.
